Exploring the Pareto front of multi-objective COVID-19 mitigation policies using reinforcement learning

### 2024-05-28 paper/policy-executions.tex
main

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsmath,bm}
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{cleveref}       % smart cross-referencing
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{doi}
\usepackage{bm}             % bold fonts for math symbols
\usepackage[misc]{ifsym}    % collection of symbols
\usepackage{caption}
\usepackage{subcaption}
\usepackage{float}          % force figure positioning in text
\usepackage{chngcntr}       % counters per section
\usepackage{tikz}
\usepackage{pgfplots,etoolbox}
\usepackage{comment}
\usetikzlibrary{external}
\usepgfplotslibrary{dateplot, groupplots, fillbetween}
\pgfplotsset{compat=newest}
\tikzexternalize[prefix=cache/]

@@

% default \cite to use parentheses
\renewcommand{\cite}[1]{\citep{#1}}

\newcommand{\foi}{\lambda}
\newcommand\todo[1]{\textcolor{red}{(TODO: #1)}}

%MDP
%state space
\newcommand{\mdpstatespace}{\mathcal{S}}
%action space
\newcommand{\mdpactionspace}{\mathcal{A}}
%mdp discount factor
\newcommand{\mdpdiscount}{\gamma}
%mdp policy
\newcommand{\mdppolicy}{\pi}
%mdpvaluefunction
\newcommand{\mdpvaluefunction}{V}
%vectorial value function
\newcommand{\momdpvaluefunction}{\mathbf{V}}
%mdpqfunction
\newcommand{\mdpqfunction}{Q}
%mdp q function estimate
\newcommand{\mdpqfunctionestimate}{\hat{Q}}
%mdp optimal q function
\newcommand{\mdpoptimalqfunction}{Q^{*}}
%q step size
\newcommand{\qstepsize}{\alpha}
%\advantageestimate
\newcommand{\advantageestimate}{\hat{A}}
%transition function
\newcommand{\mdptransition}{\mathcal{T}}
%reward fn
\newcommand{\mdprewardfn}{\mathcal{R}}
%vectorial reward fn
\newcommand{\momdprewardfn}{\bm{\mathcal{R}}}
%mdp state
\newcommand{\mdpstate}{\mathbf{s}}
%mdp reward
\newcommand{\mdpreward}{r}
\newcommand{\momdpreward}{\mathbf{r}}
\newcommand{\momdpreturn}{\mathbf{R}}
%mdp action
\newcommand{\action}{a}
%vectorial action
\newcommand{\mdpaction}{\mathbf{a}}
% the 6 social contact matrices
\newcommand{\ctotal}{C}
\newcommand{\chome}{C_{\text{home}}}
\newcommand{\cwork}{C_{\text{work}}}
\newcommand{\ctransport}{C_{\text{transport}}}
\newcommand{\cschool}{C_{\text{school}}}
\newcommand{\cleisure}{C_{\text{leisure}}}
\newcommand{\cother}{C_{\text{other}}}
% tentative momdp env name
\newcommand{\momdpname}{MOBelCov}
%other params
\newcommand{\proportionalityfactor}{q}
\newcommand{\agegroup}{k}
\newcommand{\agegroups}{K}
\newcommand{\budget}{\bm{b}}

\newcommand{\ltuple}{\langle}
\newcommand{\rtuple}{\rangle}

% color scheme for plots: plot-color-i
\definecolor{pc1}{RGB}{76, 120, 168}
\definecolor{pc2}{RGB}{245, 133, 24}
\definecolor{pc3}{RGB}{228, 87, 88}
\definecolor{pc4}{RGB}{114, 183, 178}
\definecolor{pc5}{RGB}{84, 162, 75}
\definecolor{pc6}{RGB}{238, 202, 59}
\definecolor{pc7}{RGB}{178, 121, 162}

\title{Exploring the Pareto front of multi-objective COVID-19 mitigation policies using reinforcement learning}

% Here you can change the date presented in the paper title
%\date{September 9, 1985}
% Or remove it
\date{}

\author{Mathieu Reymond \\
	Vrije Universiteit Brussel\\
	Brussels, Belgium \\
	\texttt{mathieu.reymond@vub.be} \\
	\And
	Conor F. Hayes \\
	National University of Ireland Galway\\
	Galway, Ireland\\
% 	\texttt{c.hayes13@nuigalway.ie} \\
	\And
	Lander Willem \\
	University of Antwerp \\
	Antwerp, Belgium \\
	\And
	Roxana R\u{a}dulescu \\
	Vrije Universiteit Brussel\\
	Brussels, Belgium \\
	\And
	Steven Abrams \\
	Hasselt University \\
	Hasselt, Belgium \\
	\And
	Diederik M.\ Roijers \\
	HU University of Applied Sciences Utrecht \\
	Utrecht, the Netherlands \\
	\And
	Enda Howley \\
	National University of Ireland Galway\\
	Galway, Ireland\\
	\And
	Patrick Mannion \\
	National University of Ireland Galway\\
	Galway, Ireland\\
	\And
	Niel Hens \\
	Hasselt University \\
	Hasselt, Belgium \\
	\And
	Ann Now\'{e} \\
	Vrije Universiteit Brussel\\
	Brussels, Belgium \\
	\And
	Pieter Libin \\
	Vrije Universiteit Brussel\\
	Brussels, Belgium \\
    \texttt{pieter.libin@vub.be} \\
% 	\AND
% 	Coauthor \\
% 	Affiliation \\
% 	Address \\
% 	\texttt{email} \\
}

% \author{Mathieu Reymond\inst{1}{\Letter} \and Conor F. Hayes \inst{4} \and Lander Willem  \inst{6}\and Roxana R\u{a}dulescu \inst{1} \and Steven Abrams \inst{3,7} \and Diederik M.\ Roijers\inst{1,5}\and Enda Howley\inst{4} \and Patrick Mannion\inst{4} \and Niel Hens \inst{3,6} \and Ann Now\'{e}\inst{1} \and Pieter Libin\inst{1,2,3}{\Letter}}

% \authorrunning{M. Reymond et al.}

% \institute{ 
% Artificial intelligence lab, Vrije Universiteit Brussel, Brussels, Belgium \email{\{pieter.libin,mathieu.reymond,diederik.roijers,ann.nowe\}@vub.be} 
% \and
% Rega institute, KU Leuven, Leuven, Belgium 
% \and  
% Data science institute, Hasselt University, Hasselt, Belgium
% \and
% National University of Ireland Galway, Ireland \email{\{c.hayes13\}@nuigalway.ie}
% \and
% HU University of Applied Sciences Utrecht, the Netherlands
% \and
% Centre for Health Economic Research and Modelling Infectious Diseases
% CHERMID, University of Antwerp, Belgium
% \and
% Global Health Institute, University of Antwerp, Belgium
% }

% Uncomment to override  the `A preprint' in the header
%\renewcommand{\headeright}{Technical Report}
%\renewcommand{\undertitle}{Technical Report}
\renewcommand{\shorttitle}{Exploring the Pareto front of multi-objective COVID-19 mitigation policies}

%%% Add PDF metadata to help others organize their library
%%% Once the PDF is generated, you can check the metadata with
%%% $ pdfinfo template.pdf
\hypersetup{
pdftitle={Exploring the Pareto front of multi-objective COVID-19 mitigation policies using reinforcement learning},
pdfauthor={Mathieu Reymond, Conor F.~Hayes, Lander Willem, Roxana R\u{a}dulescu, Steven Abrams , Diederik M.\ Roijers, Enda Howley, Patrick Mannion, Niel Hens, Ann Now\'{e}, Pieter Libin},
pdfkeywords={Multi-objective reinforcement learning, Epidemic control, COVID-19 epidemic models},
}

\begin{document}
\maketitle

\begin{abstract}
Infectious disease outbreaks can have a disruptive impact on public health and societal processes. As decision-making in the context of epidemic mitigation is hard, reinforcement learning provides a methodology to automatically learn prevention strategies in combination with complex epidemic models. Current research focuses on optimizing policies with respect to a single objective, such as the pathogen's attack rate. However, as the mitigation of epidemics involves distinct, and possibly conflicting, criteria (i.a., prevalence, mortality, morbidity, cost), a multi-objective decision approach is warranted to learn balanced policies. To lift this decision-making process to real-world epidemic models, we apply deep multi-objective reinforcement learning and build upon a state-of-the-art algorithm, Pareto Conditioned Networks (PCN), to learn a set of solutions that approximates the Pareto front of the decision problem. We consider the first wave of the Belgian COVID-19 epidemic, which was mitigated by a lockdown, and study different deconfinement strategies, aiming to minimize both COVID-19 cases (i.e., infections and hospitalizations) and the societal burden that is induced by the applied mitigation measures. We contribute a multi-objective Markov decision process that encapsulates the stochastic compartment model that was used to inform policymakers during the COVID-19 epidemic. As these social mitigation measures are implemented in a continuous action space that modulates the contact matrix of the age-structured epidemic model, we extend PCN to this setting. We evaluate the solution set that PCN returns, and observe that it correctly learns to reduce the social burden whenever the hospitalization rates are sufficiently low. In this work, we thus demonstrate that multi-objective reinforcement learning is attainable in complex epidemiological models and provides essential insights to balance complex mitigation policies.
\end{abstract}

% keywords can be removed
\keywords{Multi-objective reinforcement learning  \and Epidemic control \and COVID-19 epidemic models}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Infectious disease outbreaks represent a major global challenge \cite{miranda2022tale}. To this end, understanding the complex dynamics that underlie these epidemics is essential. Epidemiological transmission models allow us to capture and understand such dynamics and facilitate the study of prevention strategies through simulation. However, developing efficient mitigation strategies remains a challenging process, given the non-linear and complex nature of epidemics.

To address these challenges, reinforcement learning provides a methodology to automatically learn mitigation strategies in combination with complex epidemic models~\cite{libin2020}.
Previous research focused on optimizing policies with respect to a single objective, such as the pathogen's attack rate, while the mitigation of epidemics is a problem that inherently covers distinct  and possibly conflicting criteria (i.a., prevalence, mortality, morbidity, cost). Therefore, optimizing on a single objective requires that  these distinct criteria are somehow aggregated into a single metric.
Typically, a decision maker will rely on an expert's assistance to manually design such a metric. However, this is a challenging process, as the decision maker is generally not aware of all possible trade-offs, and as such this may result in a metric that exhibits suboptimal performance. Furthermore, manually designing such metrics is time-consuming, costly and error-prone, as this non-intuitive process requires repetitive and tedious tuning to achieve the desired behavior~\cite{roijers2013survey}. Moreover, taking a single objective approach has several other limiting factors. For example, it reduces the explainability of the learned solution, as we cannot compare the learned behavior with alternatives. Moreover, the preferences of the decision maker might change over time, in which case the optimization process needs to be done again~\cite{hayes2021practical}.

This challenging process can be circumvented by taking an explicit multi-objective approach that aims to learn the different trade-offs regarding the considered criteria. By assuming that a decision maker will always prefer solutions for which at least one objective improves, it is possible to learn a set of optimal solutions referred to as the \emph{Pareto front}~\cite{hayes2021practical}. This enables decision makers to review each solution on the Pareto front before making a decision, thereby being aware of the trade-offs that each solution implies.

In this work, we investigate the use of \emph{multi-objective reinforcement learning} (MORL) to learn a set of solutions that approximate the Pareto front of multi-objective epidemic mitigation strategies. We consider the first wave of the Belgian COVID-19 epidemic, which was mitigated by a strict lockdown~\cite{willem2021impact}. When the incidence of confirmed cases was steadily decreasing, epidemiological experts were tasked to investigate deconfinement strategies, to reduce the severe social contact and mobility restrictions.
Here, we consider an epidemiological model that was constructed to describe the Belgian COVID-19 epidemic and was fitted to hospitalization incidence data and serial sero-prevalence data~\cite{abrams2021modelling}.
This model constitutes a stochastic discrete-time age-structured compartmental model that simulates mitigation strategies by varying social distancing parameters concerning school, work and leisure contacts.  
Based on this model, we contribute a novel multi-objective epidemiological reinforcement learning environment~(\momdpname), in the form of a multi-objective Markov decision process (MOMDP) \cite{roijers2013survey}. \momdpname\ encapsulates the epidemiological model developed by \citet{abrams2021modelling} to implement state transitions, with an action space that combines a proportional reduction of school, work and leisure contacts at each time step, Furthermore, it defines a reward function based on two objectives: the attack rate (i.e., proportion of the population affected by the pathogen) and the social burden that is induced by the mitigation measures.

To learn and explore the trade-offs between the attack rate and social burden we use a state-of-the-art MORL approach based on Pareto Conditioned Networks (PCN)~\cite{reymond2022pcn}. PCN uses a single neural network to learn the policies that belong to the Pareto front.
As PCN is an algorithm designed for discrete action-spaces, we extend it towards continuous action-spaces to accommodate \momdpname's action-space. With this continuous action variant of PCN, we explore the Pareto front of multi-objective COVID-19 mitigation policies.

By evaluating the solution set of mitigation policies learned by PCN, we observe that PCN minimizes the social burden in scenarios where hospitalization rates are sufficiently low. Therefore, in this work we illustrate that multi-objective reinforcement learning can provide important insights concerning the trade-offs between complex mitigation polices in real-world epidemiological models.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{COVID-19 model and \momdpname\  MOMDP}
\label{sec:sars-cov2-momdp}

\emph{Reinforcement learning} (RL) is a set of machine learning techniques to learn the optimal behavior for sequential decision problems. Through trial and error, the RL agent interacts with its environment and observes the consequences of its decisions on the environment. Decisions that led to desirable consequences are reinforced, such that this behavior can be repeated. Otherwise, the agent explores the environment by trying different decisions. Developing mitigation strategies for epidemic control is a sequential decision problem on which we can apply RL, which will learn a desirable strategy.

In this Section, we first introduce the compartment model that we use to simulate the epidemic (Sec.~\ref{sec:sars-cov2-momdp_model}). Then, we explain the intervention strategies that can be made to control the evolution of the epidemic (Sec.~\ref{sec:interventions}). Afterwards, we introduce the multi-objective reinforcement learning (MORL) framework (Sec.~\ref{sec:reinforcement-learning}). Finally, we propose \momdpname, a MORL environment that encompasses the epidemiological model and intervention strategies, so we can learn optimal mitigation strategies (Sec.~\ref{sec:covid-momdp}).

In the following, we use bold notation to denote variables with a vectorial value, while non-bold notation for its scalar counterpart.
\subsection{Stochastic compartment model for SARS-CoV-2}
\label{sec:sars-cov2-momdp_model}
To evaluate non-pharmaceutical interventions, we consider the compartmental model presented by Abrams et al., that was used to investigate exit strategies in Belgium after the first epidemic wave of SARS-CoV-2~\cite{abrams2021modelling}. This model concerns a discrete-time stochastic model, that considers an age-structured population. The model generalizes a standard SEIR model\footnote{A standard SEIR model divides the population into four different states, i.e., susceptible, exposed, infectious and recovered individuals.}, extended to capture the different stages of disease spread and history that are associated with SARS-CoV-2 (i.e., pre-symptomatic, asymptomatic, symptomatic with mild symptoms and symptomatic with severe symptoms) and to represent the stages associated with severe disease, i.e., hospitalization, admission to the intensive care unit (ICU) and death.

A visual representation of the model is depicted in Fig.~\ref{fig:compartment_model}. It shows the different compartments, as well as the flow rates at which individuals move between compartments.

\begin{figure}[t!]
    \centering
    \begin{tikzpicture}
     \draw (0.0pt, 0.0pt)node[fill=green!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](0){$\textbf{S}$};
     \draw (60.0pt, 0.0pt)node[fill=orange!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](1){$\textbf{E}$};
     \draw (120.0pt, 0.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](2){$\textbf{I}^{presym}$};
     \draw (180.0pt, 10.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](3){$\textbf{I}^{asym}$};
     \draw (170.0pt, -30.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](4){$\textbf{I}^{mild}$};
     \draw (170.0pt, -70.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](5){$\textbf{I}^{sev}$};
     \draw (230.0pt, -50.0pt)node[fill=blue!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](6){$\textbf{I}^{hosp}$};
      \draw (75.0pt, -70.0pt)node[fill=gray!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](10){$\textbf{H}^{new}$};
     \draw (230.0pt, -90.0pt)node[fill=blue!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](7){$\textbf{I}^{icu}$};
     \draw (300.0pt, -50.0pt)node[fill=orange!50, thick, minimum height=0.6cm,minimum width=0.6cm, draw](8){$\textbf{R}$};
     \draw (300.0pt, -90.0pt)node[fill=purple!50, thick, minimum height=0.6cm,minimum width=0.6cm, draw](9){$\textbf{D}$};

     %\path[->] (6) edge[dotted, bend right = 50]
     \path[->] (0) edge[] node[above]{$\boldsymbol{\lambda}(t)$} (1);
     \path[->] (1) edge[] node[above]{$\gamma$} (2);
     \path[->] (2) edge[] node[below]{$\textbf{p}\theta$} (3);
     \path[->] (2) edge[] node[near end, below left]{$(1-\textbf{p})\theta$} (4);
     \path[->] (4) edge[] node[left]{$\boldsymbol{\psi}$} (5);
     \path[->] (5) edge[] node[near end, below left]{$(1-\phi_{1})\boldsymbol{\omega}$} (7);
     \path[->] (5) edge[] node[above]{$\phi_{1}\boldsymbol{\omega}$} (6);
     \path[->] (7) edge[] node[near end, below right]{$\boldsymbol{\delta}_{4}$} (8);
     \path[->] (7) edge[] node[above]{$\boldsymbol{\tau}_{2}$} (9);
     \path[->] (6) edge[] node[above]{$\boldsymbol{\delta}_{3}$} (8);
     \path[->] (6) edge[] node[near start, below left]{$\boldsymbol{\tau}_{1}$} (9);
     \path[->] (3) edge[bend left = 30] node[above right]{$\delta_{1}$} (8);
     \path[->] (4) edge[bend left = 30] node[above left]{$\boldsymbol{\delta}_{2}$} (8);
      \path[->] (5) edge[] node[below]{$\textbf{I}^{sev} - \textbf{H}^{new}$} (10);
    \end{tikzpicture}

    \caption{Schematic diagram of the compartmental model for SARS-CoV-2 presented by \citet{abrams2021modelling} which is used to derive the MOMDP.}
    \label{fig:compartment_model}
\end{figure}

These flow rates are defined by a set of ordinary differential equations, which are outlined as follows:

\begin{align*}
\frac{d\textbf{S}(t)}{dt} & = -\lambda(t)(\textbf{S})(t), \\
%
\frac{d\textbf{E}(t)}{dt} & = \lambda(t)(\textbf{S})(t) - \gamma \textbf{E}(t), \\
%
\frac{d\textbf{I}^{presym}(t)}{dt} & = \gamma \textbf{E}(t) - \theta \textbf{I}^{presym}(t), \\
%
\frac{d\textbf{I}^{asym}(t)}{dt} & = \theta p \textbf{I}^{presym}(t) - \delta_{1} \textbf{I}^{asym}(t), \\
%
\frac{d\textbf{I}^{mild}(t)}{dt} & = \theta (1 - p) \textbf{I}^{presym}(t) - \{ \psi + \omega_{2} \} \textbf{I}^{mild}(t), \\
%
\frac{d\textbf{I}^{sev}(t)}{dt} & = \psi \textbf{I}^{mild}(t) - \omega \textbf{I}^{sev}(t), \\
%
\frac{d\textbf{I}^{hosp}(t)}{dt} & = \phi_{1} \omega \textbf{I}^{sev}(t) - \{ \delta_{3} + \tau_{1} \} \textbf{I}^{hosp}(t), \\
%
\frac{d\textbf{I}^{icu}(t)}{dt} & = (1 - \phi_1) \omega \textbf{I}^{sev}(t) - \{ \delta_{4} + \tau_{2} \} \textbf{I}^{icu}(t), \\
%
\frac{d\textbf{D}(t)}{dt} & = \tau_{1} \textbf{I}^{hosp}(t) - \tau_{2} \textbf{I}^{icu}(t), \\
%
\frac{d\textbf{R}(t)}{dt} & = \delta_{1} \textbf{I}^{asym}(t) + \delta_{2} \textbf{I}^{mild}(t) + \delta_{3} \textbf{I}^{hosp}(t) + \delta_{4} \textbf{I}^{icu}(t)
\end{align*}
%

In this set of ordinary differential equations, each state variable represents a vector over all age groups for a particular compartment at time $t$. For example, $\textbf{S} = (S_{1}(t), S_{2}(t), ..., S_{\agegroup}(t))^{T}$ is the vector representing the susceptible members of the population of each age group $\agegroup$ at time $t$. Infection dynamics are governed by an age-specific force of infection $\foi$:

\begin{equation}
    \foi(\agegroup,t) = \sum_{\agegroup'=1}^{\agegroups}\beta(\agegroup,\agegroup')I_{\agegroup'}(t),
\end{equation}
%
where $\agegroups$ is the total number of age groups, and $\beta(\agegroup,\agegroup')$ is the time-invariant transmission rate that encodes the average per capita rate at which an infectious individual in age group $\agegroup$ makes an effective contact with a susceptible individual in age group $\agegroup'$, per unit of time.

As we consider an age-structured population, we consider this extended SEIR structure for $\agegroups = 10$ age groups,  i.e., $[0-10), [10-20), [20-30), [30-40), [40-50), [50-60), [60-70), [70-80), [80-90), [90,100+)$. Contacts of the different age-groups, which impact the propagation rate of the epidemic, are modeled using social contact matrices. We define a social contact matrix for 6 different social environments: $\chome, \cwork, \ctransport, \cschool, \cleisure, \cother$, for the home, work, transport, school, leisure, other environments respectively. The social contact matrix across all social environments is defined as:

\begin{equation}
    \ctotal = \chome + \cwork + \ctransport + \cschool + \cleisure + \cother
\end{equation}

Under the social contact hypothesis \cite{wallinga2006using}, we have that:
%
\begin{equation}
\beta(\agegroup,\agegroup') = \proportionalityfactor \cdot \ctotal(\agegroup,\agegroup'),
\end{equation}
%
where $\proportionalityfactor$ is a proportionality factor.

Following \citet{abrams2021modelling}, we rely on distinct social contact matrices for symptomatic and asymptomatic individuals, respectively $\ctotal_{s}$ and $\ctotal_{a}$. Therefore, we define the transmission rates for both symptomatic and asymptomatic individuals as follows:

\begin{equation}
\boldsymbol{\beta}_{s}(\agegroup, \agegroup) = \proportionalityfactor_{s} \cdot C_{s}(\agegroup, \agegroup'),    
\end{equation}
and
\begin{equation}
\boldsymbol{\beta}_{a}(\agegroup, \agegroup') = \proportionalityfactor_{a} \cdot C_{a}(\agegroup, \agegroup').  
\end{equation}

The age-dependent force of infection can be defined as follows:

\begin{equation}
    \boldsymbol{\lambda}(t) = \boldsymbol{\beta}_{a} \times \{ \textbf{I}^{presym}(t) + \textbf{I}^{asym}(t) \} +
    \boldsymbol{\beta}_{s} \times \{ \textbf{I}^{mild}(t) + \textbf{I}^{sev}(t) \},
\end{equation}
%
where $\boldsymbol{\lambda}(t) = (\lambda(1, t), \lambda(2, t), ..., \lambda(\agegroups, t))$. For all further information about the different compartments and parameters we refer the reader to the work of \citet{abrams2021modelling}.

\todo{epi friends, please check} Time- and age-dependent behavioral changes introduce substantial uncertainty regarding the outcome of an outbreak. To evaluate intervention strategies that modulate such behavioral changes, the use of stochastic epidemiological models is warranted \cite{abrams2021modelling}. Moreover, the effect of stochasticity is most pronounced when studying the initial growth of an epidemic \cite{britton2009epidemic}, be it during its emergence, or when implementing deconfinement strategies.

By formulating the set of differential equations defined above, as a chain-binomial, we can obtain stochastic trajectories from this model \cite{bailey1975mathematical}. A chain-binomial model assumes a stochastic model where infected individuals are generated by some underlying probability distribution. For this stochastic model we consider a time interval $(t, t +h]$, where $h$ is defined as the length between two consecutive time points. In this work we set $h = \frac{1}{240}$.

\subsection{Interventions strategies}
\label{sec:interventions}
To model different types of interventions, we follow the contact reduction scheme presented by \citet{abrams2021modelling}. Firstly, to consider distinct exit scenarios, we modulate the social contact matrices to reflect a contact reduction in a particular age group.

We consider a contact reduction function that imposes a proportional reduction of work (including transport) $p_w$, school $p_s$ and leisure $p_l$ contacts, which is implemented as a linear combination of social contact matrices:
\begin{equation}
\label{eq:scm}
    \hat{C}(p_w, p_s, p_l) = \chome + p_w (\cwork + \ctransport) + p_s \cschool + p_l (\cleisure + \cother)
\end{equation}

We denote $\hat{C}_t$ the social contact matrix at timestep $t$, resulting from the reduction function $\hat{C}$. Secondly, we assume that compliance to the interventions is gradual and model this using a logistic compliance function (see details in Sec.~\ref{sec:compliance-function} of the Appendix).

\subsection{Multi-Objective Reinforcement Learning}
\label{sec:reinforcement-learning}
Real-world decisions problems typically consider multiple and possibly conflicting objectives. Multi-objective reinforcement learning (MORL) can be used to find optimal solutions for sequential decision making problems with multiple objectives~\cite{hayes2021practical}. Multi-objective sequential problems are typically modelled as a multi-objective Markov decision process (MOMDP), i.e., a tuple, $\mathcal{M} = \ltuple\mdpstatespace, \mdpactionspace, \mdptransition, \mdpdiscount, \momdprewardfn\rtuple$, where $\mdpstatespace$, $\mdpactionspace$ denote the state and action spaces respectively, $\mdptransition \colon \mdpstatespace \times \mdpactionspace \times \mdpstatespace  \to \left[ 0, 1 \right]$ denotes a probabilistic transition function, $\mdpdiscount$ is a discount factor determining the importance of future rewards and $\momdprewardfn \colon \mdpstatespace \times \mdpactionspace \times \mdpstatespace \to \mathbb{R}^n$ is an $n$-dimensional vector-valued immediate reward function, where $n$ corresponds to the number of objectives.

While MOMDPs define the actions an agent can take, it does not define the agent's behavior, i.e., which actions are taken in each state. Given this MOMDP, an agent follows a policy $\mdppolicy$, that expresses the probability to take action $\action \in \mdpactionspace$ when in state $\mdpstate \in \mdpstatespace$: $\mdpstatespace \times \mdpactionspace \to \left[ 0, 1 \right]$. We measure the performance of $\mdppolicy$ through the expected sum of discounted rewards (denoted the \emph{Value}) it achieves from the initial state $\mdpstate_0$ until the end of the decision problem:

\begin{equation}
    \momdpvaluefunction^\mdppolicy = \mathbb{E} \left [ \sum_{t=0}^h \mdpdiscount^t \momdpreward_t |\ \mdppolicy, \mdpstate_0 \right ],
\end{equation}
%
where $\momdpreward_t$ corresponds to the multi-objective reward observed at time $t$, by following policy $\mdppolicy$, given by the reward function $\mdprewardfn(\mdpstate_t, \action_t, \mdpstate_{t+1})$, where $\mdpstate_t$ is the current state, $\action_t$ corresponds to the action that was taken, and $\mdpstate_{t+1}$ signifies the state reached by taking action $\action_t$.

On the one hand, for single-objective RL, where $n = 1$, the goal is to find the policy $\mdppolicy^{*}$ that maximizes the $\mdpvaluefunction$-value:
\begin{equation}
\mdppolicy^*= \arg\max_\mdppolicy \mdpvaluefunction^\mdppolicy.
\end{equation}

On the other hand, for MORL, $n > 1$ which leads to vectorial returns. In this case, there can be policies for which, without any additional information, it is impossible to know if one is better than the other. For example, it is impossible to decide which policy between $\mdppolicy_1, \mdppolicy_2$ is optimal if both policies lead to expected returns $\momdpvaluefunction^{\mdppolicy_1}=(0,1), \momdpvaluefunction^{\mdppolicy_2}=(1,0)$ respectively. We call these solutions \emph{non-dominated}, i.e., solutions for which it is impossible to improve an objective without hampering another. The set that contains all the non-dominated solutions of the decision problem is called the \emph{Pareto front} $\mathcal{F}$. Our goal is to find the set of policies that lead to all the $\mdpvaluefunction$-values contained in the Pareto front $\Pi^* = \{\ \pi | \momdpvaluefunction^{\pi} \in \mathcal{F}\}$. In general, we call any set of $\mdpvaluefunction$-values a \emph{solution set}. When a solution set contains only  non-dominated $\mdpvaluefunction$-values, it is referred to as a \emph{coverage set}. In the case that no $\mdppolicy$ exists that has a $\momdpvaluefunction^{\pi}$ dominating any of the solutions in a coverage set, then this coverage set is the Pareto front.

\subsubsection{Multi-Objective Metrics}
Comparing the learned coverage sets produced by different algorithms is a non-trivial task, as one algorithm's output might dominate the other in some region of the objective-space, but be dominated in another. Intuitively, one would generally prefer the algorithm that covers a wider range of decision maker preferences. %In this work we use several metrics to evaluate our algorithm's performance.

A widely used metric in the literature is called the \emph{hypervolume}~\cite{zitzler2003}. This metric evaluates the learned coverage set by computing its volume with respect to a fixed reference point. The reference point is taken as a lower bound on the achievable returns so that the volumes are always positive. By definition, the solutions contained in the Pareto front dominate all other possible solutions. Thus, no other solution can further increase the volume under the Pareto front. This means that the hypervolume is the highest for the Pareto front. One drawback of the hypervolume is that it can be difficult to interpret. For example, when working in high-dimensional objective-spaces, adding or removing a single point can drastically change hypervolume values, especially if the point lies close to an extremum of said space.

To counterpoise these shortcomings, we evaluate our work using an additional metric called the $\varepsilon$-indicator $I_\varepsilon$~\cite{zitzler2003}, which measures how close a coverage set is to the Pareto front $\mathcal{F}$. Intuitively, $I_\varepsilon$ shows that any solution of $\mathcal{F}$ is \emph{at most} $\varepsilon$ better with respect to each objective $o$ than the closest solution of the evaluated coverage set:

\begin{equation}
\label{eq:epsilon-metric}
    I_{\varepsilon} = 
    \inf_{\varepsilon\in\mathbb{R}} 
    \{ 
    \forall\ \momdpvaluefunction^\pi\!{\in}\ \mathcal{F}, ~
    \exists\ \momdpvaluefunction^{\pi'}\!{\in}\ \hat{\Pi} :\ 
    || \momdpvaluefunction^\pi - \momdpvaluefunction^{\pi'} ||_{\infty} \le \varepsilon
    \},
\end{equation}
%
where $||.||_{\infty}$, the L-infinity norm, is defined as the magnitude of the largest entry of a vector.

The main disadvantage of this metric is that we need the true Pareto front to compute it, which is unknown for our MOMDP (see Sec.~\ref{sec:sars-cov2-momdp}). To still gain insights from our learned policies, we approximate the true Pareto front using the non-dominated policies across all runs.

>>>>

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsmath,bm}
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{cleveref}       % smart cross-referencing
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{doi}
\usepackage{bm}             % bold fonts for math symbols
\usepackage[misc]{ifsym}    % collection of symbols
\usepackage{caption}
\usepackage{subcaption}
\usepackage{float}          % force figure positioning in text
\usepackage{chngcntr}       % counters per section
\usepackage{tikz}
\usepackage{pgfplots,etoolbox}
\usepackage{comment}
\usetikzlibrary{external}
\usepgfplotslibrary{dateplot, groupplots, fillbetween}
\pgfplotsset{compat=newest}
\tikzexternalize[prefix=cache-policies/]

\newcommand{\momdpname}{MOBelCov}

@@

\pagestyle{empty}
\begin{document}
\section{Policy executions}

Depending on the budget, PCN learns a coverage set containing more than 150 different policies. To gain a better insight about their behavior, and how they differ from each other, we plot executions of each learned policy in Fig.~\ref{fig:policy-executions-b2-0}-\ref{fig:policy-executions-b2-140}. The plots are displayed from the least restrictive policy in terms of social burden to the most restrictive one. Since the \momdpname\ model is stochastic, we show 10 executions of the same policy, on each plot.

\foreach \i/\j in {0/19, 20/39, 40/59, 60/79, 80/99, 100/119, 120/139, 140/153}{
    \begin{figure}
        \centering
        \plotPolicies{\i}{\j}{2}
        \caption{Execution of policies \i\ to \j, with a budget of 2.}
        \label{fig:policy-executions-b2-\i}
    \end{figure}
}

\foreach \i/\j in {0/19, 20/39, 40/59, 60/79, 80/99, 100/119, 120/139, 140/159, 160/160}{
    \begin{figure}
        \centering
        \plotPolicies{\i}{\j}{3}
        \caption{Execution of policies \i\ to \j, with a budget of 3.}
        \label{fig:policy-executions-b3-\i}
    \end{figure}
}

\foreach \i/\j in {0/19, 20/39, 40/59, 60/79, 80/93}{
    \begin{figure}
        \centering
        \plotPolicies{\i}{\j}{4}
        \caption{Execution of policies \i\ to \j, with a budget of 4.}
        \label{fig:policy-executions-b4-\i}
    \end{figure}
}

\foreach \i/\j in {0/19, 20/39, 40/59, 60/79, 80/99, 100/114}{
    \begin{figure}
        \centering
        \plotPolicies{\i}{\j}{5}
        \caption{Execution of policies \i\ to \j, with a budget of 5.}
        \label{fig:policy-executions-b5-\i}
    \end{figure}
}

\foreach \i/\j in {0/19, 20/39, 40/59, 60/79, 80/99, 100/116}{
    \begin{figure}
        \centering
        \plotPolicies{\i}{\j}{None}
        \caption{Execution of policies \i\ to \j, with no budget restriction ($b=\infty$).}
        \label{fig:policy-executions-binf-\i}
    \end{figure}
}

\end{document}
